---
title: "MRSA Mouse Liver"
output: html_document
date: "2024-10-28"
---

```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
# install.packages("BiocManager")
# BiocManager::install(version = "3.14")

# install.packages("devtools")
# devtools::install_github('cole-trapnell-lab/monocle3')

library(monocle3)

library(ggplot2)
library(Seurat)
# library(DESeq2)

# library(scales)
# library(ggpubr)
# library(clusterProfiler)
# library(openxlsx)
# library(org.Hs.eg.db)
# library(org.Mm.eg.db)
# library(ReactomePA)
# library(ComplexHeatmap)
# library(ComplexUpset)
# library(ggvenn)
# library(jsonlite)

path = 'C:/Users/kmccr/Desktop/MRSA Mouse Kidney for Takashi/'
```

### Import Seurat Objects
```{r}
A2 = readRDS(paste0(path,'Data/seurat_objects/achromoday2_seurat.RDS'))
A4 = readRDS(paste0(path,'Data/seurat_objects/achromoday4_seurat.RDS'))
S2 = readRDS(paste0(path,'Data/seurat_objects/standardday2_seurat.RDS'))
S4 = readRDS(paste0(path,'Data/seurat_objects/standardday4_seurat.RDS'))
A2[["slice1.2"]] <- NULL
A4[["slice1.2"]] <- NULL
S2[["slice1"]] <- NULL
S4[["slice1"]] <- NULL
# kidneymerge = readRDS(paste0(path,'kidneyMerge_annotated.RDS'))

samp_dict <- list("Day 2 - Achromo" = A2, "Day 4 - Achromo" = A4, "Day 2 - Standard" = S2, "Day 4 - Standard" = S4)

sample = S4
sample_description = 'Day 4 - Standard'
min_expression_filter = 1
```

### Define function
```{r}
library(Matrix)

gene_annotate = function(expression_matrix){
  # Step 1: Extract gene IDs (row names of the matrix)
  gene_ids <- rownames(expression_matrix)
  
  # Step 2: Create or extract gene short names
  gene_short_names <- gene_ids
  
  # Step 3: Calculate num_cells_expressed
  num_cells_expressed <- rowSums(expression_matrix > 0)
  
  # Step 4: Create the gene annotation matrix
  gene_annotation <- data.frame(
    id = gene_ids,
    gene_short_name = gene_short_names,
    num_cells_expressed = num_cells_expressed
  )
  
  return(gene_annotation)
}
```

### SELECT SAMPLE AND LOAD
```{r}
sample = A4

# expression_matrix1 <- readRDS(url("https://depts.washington.edu:/trapnell-lab/software/monocle3/celegans/data/packer_embryo_expression.rds"))
# cell_metadata1 <- readRDS(url("https://depts.washington.edu:/trapnell-lab/software/monocle3/celegans/data/packer_embryo_colData.rds"))
# gene_annotation1 <- readRDS(url("https://depts.washington.edu:/trapnell-lab/software/monocle3/celegans/data/packer_embryo_rowData.rds"))

expression_matrix = GetAssayData(sample, assay = "SCT", layer = "counts")
cell_metadata = as.data.frame(sample@meta.data)
cell_metadata$MRSA = ifelse(cell_metadata$MRSAlevel > 1, "MRSA-Yes", "MRSA-No")
gene_annotation = gene_annotate(expression_matrix)

cds_raw <- new_cell_data_set(expression_matrix, cell_metadata = cell_metadata, gene_metadata = gene_annotation)
```

### Pre-process, reduce dimensionality
```{r}
# preprocess
# Note: Your data will not have the loading batch information demonstrated here, you will correct batch using your own batch information.
cds <- preprocess_cds(cds_raw, num_dim = 50)
plot_pc_variance_explained(cds)

# batch effect correction
# cds <- align_cds(cds, alignment_group = "batch", residual_model_formula_str = "~ bg.300.loading + bg.400.loading + bg.500.1.loading + bg.500.2.loading + bg.r17.loading + bg.b01.loading + bg.b02.loading")

# reduce dim
cds <- reduce_dimension(cds)
```

### Visualize
```{r}
# visualize
plot_cells(cds, label_groups_by_cluster=FALSE,  color_cells_by = "MRSA")
plot_cells(cds, label_groups_by_cluster=FALSE,  color_cells_by = "labels")
```

### Cluster
```{r}
cds <- cluster_cells(cds)
plot_cells(cds, color_cells_by = "partition")
```

### Learn trajectory graph
```{r}
cds <- learn_graph(cds)
p1 = plot_cells(cds,
           color_cells_by = "labels",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)

p1

# ggsave("S2 trajectory.png", plot = p, width = 8, height = 6)
```


# Choose root nodes and plot
# RUN CODE ONLY FOR SELECTED SAMPLE (S2,S4,A2,A4)!

# S2
## S2 PSEUDOTIME -- PT-S3 and principal-like
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S2 PSEUDOTIME -- PT-S3 and principal-like',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S2 PSEUDOTIME -- PT-S3 and principal-like.png", combined_plot, width = 15, height = 15, dpi = 300)
```

## S2 PSEUDOTIME -- Endo
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S2 PSEUDOTIME -- Endo',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S2 PSEUDOTIME -- Endo.png", combined_plot, width = 15, height = 15, dpi = 300)
```

## S2 PSEUDOTIME -- Injured
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S2 PSEUDOTIME -- Injured',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S2 PSEUDOTIME -- Injured.png", combined_plot, width = 15, height = 15, dpi = 300)
```


# S4
## S4 PSEUDOTIME -- myeloid and neutrophil
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S4 PSEUDOTIME -- myeloid and neutrophil',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S4 PSEUDOTIME -- myeloid and neutrophil.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## S4 PSEUDOTIME -- ureteral
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S4 PSEUDOTIME -- ureteral',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S4 PSEUDOTIME -- ureteral.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## S4 PSEUDOTIME -- other
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'S4 PSEUDOTIME -- other',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("S4 PSEUDOTIME -- other.png", combined_plot, width = 15, height = 15, dpi = 300)
```


# A2
## A2 PSEUDOTIME -- endo
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A2 PSEUDOTIME -- endo',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A2 PSEUDOTIME -- endo.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## A2 PSEUDOTIME -- other
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A2 PSEUDOTIME -- other',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A2 PSEUDOTIME -- other.png", combined_plot, width = 15, height = 15, dpi = 300)
```


#A4
## A4 PSEUDOTIME -- 1
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A4 PSEUDOTIME -- 1',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A4 PSEUDOTIME -- 1.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## A4 PSEUDOTIME -- 2
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A4 PSEUDOTIME -- 2',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A4 PSEUDOTIME -- 2.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## A4 PSEUDOTIME -- 3
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A4 PSEUDOTIME -- 3',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A4 PSEUDOTIME -- 3.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## A4 PSEUDOTIME -- 4
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A4 PSEUDOTIME -- 4',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A4 PSEUDOTIME -- 4.png", combined_plot, width = 15, height = 15, dpi = 300)
```
## A4 PSEUDOTIME -- 5
```{r}
library(patchwork)
cds <- order_cells(cds)

p2 = plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=1.5)

sample$pseudotime = cds@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]
sample$pseudotime[is.infinite(sample$pseudotime)] <- 0
sample$pseudotime <- as.numeric(sample$pseudotime)

p3 = SpatialDimPlot(sample, group.by = 'MRSAlevel', interactive = FALSE)
p4 = SpatialFeaturePlot(sample, 'pseudotime', interactive = FALSE)

combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = 'A4 PSEUDOTIME -- 5',
    theme = theme(plot.title = element_text(size = 30))
  )

ggsave("A4 PSEUDOTIME -- 5.png", combined_plot, width = 15, height = 15, dpi = 300)
```







